---
title: "hatecrime"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Hate Crimes -- Group Project
Devon Maloney 
1/29/2020

To set up the data, I seperated the national individual records of data and place into counts of incidents per month in excell by using the COUNTIF function for all unique month + year combinations. I did this again for just the New Jersery data, a state with a more complete record of hate crime data collection. 
-----
## Packages and Directory
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
install.packages("ggplot2") 
install.packages("RColorBrewer")
install.packages("ggridges")

library(ggplot2)
library(RColorBrewer) 
library(ggridges)
library(readr)

setwd ("~/Documents/UNC/Spring 2020/456 Geovisualizing Change/HateCrime")

hatedata <- read_csv("monthyear_indicident.csv")
hatedataNJ <- read_csv("monthyear_indicidentNJ.csv")
```

## Bar graphs for seasonality 
Make a bar graph with Title, axis labels and caption with simple ggplot bargraph. 

```{r}
p = ggplot(hatedata, aes(x=Month, y = incident)) + geom_bar(stat="identity") 
pnj = ggplot(hatedataNJ, aes(x=Month, y = incident)) + geom_bar(stat="identity") 

#Add labels
p = p + labs(title="Seaonality of hate crimes", x ="Month", y = "Number of indicdents", subtitle = "National rates", caption = "(based on data from the FBI)") 
p 

pnj = pnj + labs(title="Seaonality of hate crimes", x ="Month", y = "Number of indicdents", subtitle = "New Jersey rates", caption = "(based on data from the FBI)") 
pnj 
```

## Warming cirlces 

```{r}
# National  
# Colors 
p = p + aes(fill = incident) # assign gradient of blue color to the temp value 
p = ggplot(hatedata, aes(x=Monthyear, y = incident, fill = incident)) + geom_bar(stat="identity")
display.brewer.all() # look at he brewer color library 
p + scale_fill_gradientn(colours = colorRampPalette(brewer.pal(11,"RdYlBu"))(11)) # red to blue  
p <- p + scale_fill_gradientn(colours = colorRampPalette(rev(brewer.pal(11,"RdYlBu")))(11)) # REV to adjust scale from blue to red 
p
p + aes(y = 300) # make it warming stripes
p + aes(x = as.factor(Monthyear), y = 300) # to avoid missing stripes, make yearmonth a factor 

# Add add minial theme, legend, titles, axis labels. 
# Ideally, this section would also include a way to label the year on the x axis, but I was unable to do that. 
p + aes(x = as.factor(Monthyear), y = 300) + theme_minimal() + theme(axis.text = element_blank(), axis.title = element_text(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position='right') + labs (x= "Month", y = "Incidents", title = "National Incidents of Hate Crime", subtitle = "Monthly from 1991 - 2017", caption = "(based on data from the FBI)")

# New Jersey 
# Colors 
pnj = pnj + aes(fill = incident) # assign gradient of blue color to the temp value 
pnj = ggplot(hatedataNJ, aes(x=Monthyear, y = incident, fill = incident)) + geom_bar(stat="identity")
display.brewer.all() # look at the brewer color library, select package 
pnj + scale_fill_gradientn(colours = colorRampPalette(brewer.pal(11,"RdYlBu"))(11)) # red to blue  
pnj <- pnj + scale_fill_gradientn(colours = colorRampPalette(rev(brewer.pal(11,"RdYlBu")))(11)) # REV to adjust scale from blue to red 
pnj + aes(y = 300) # make it warming stripes
pnj + aes(x = as.factor(Monthyear), y = 300) # to avoid missing stripes, make yearmonth a factor 

# Add add minial theme, legend, titles, axis labels. 
# Ideally, this section would also include a way to label the year on the x axis, but I was unable to do that. 
pnj + aes(x = as.factor(Monthyear), y = 300) + theme_minimal() + theme(axis.text = element_blank(), axis.title = element_text(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position='right') + labs (x= "Month", y = "Incidents", title = "New Jersey Incidents of Hate Crime", subtitle = "Monthly from 1991 - 2017", caption = "(based on data from the FBI)")


```

## PETAL CIRCLE 
For this section, the independent variable is swtiched from Monthyear to Month. 

Tried to remove the unnessicary axises. Unknown why national graph is on 0.5 month increments and why the NJ is correctly on 1 month incriments. Add add minimal theme, legend, titles, axis labels. 


```{r}
# National  
# Colors 
p = p + aes(fill = incident) # assign gradient of blue color to the temp value 
p = ggplot(hatedata, aes(x=Month, y = incident, fill = incident)) + geom_bar(stat="identity")
display.brewer.all() # look at he brewer color library 
p + scale_fill_gradientn(colours = colorRampPalette(brewer.pal(11,"RdYlBu"))(11)) # red to blue  
p <- p + scale_fill_gradientn(colours = colorRampPalette(rev(brewer.pal(11,"RdYlBu")))(11)) # REV to adjust scale from blue to red 
p
p + aes(y = 300) # make it warming stripes
p + aes(x = as.factor(Month), y = 300) # to avoid missing stripes, make month a factor 

#Add labels 
# Ideally, this section would also include a way to label the year on the x axis. 
p + aes(x = as.factor(Month), y = 300) + theme_minimal() + theme(axis.text = element_blank(), axis.title = element_text(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position='right') + labs (x= "Month", y = "Incidents", title = "National Incidents of Hate Crime", subtitle = "Monthly from 1991 - 2017", caption = "(based on data from the FBI)")

# New Jersey 
# Colors 
pnj = pnj + aes(fill = incident) # assign gradient of blue color to the temp value 
pnj = ggplot(hatedataNJ, aes(x=Month, y = incident, fill = incident)) + geom_bar(stat="identity")
display.brewer.all() # look at he brewer color library 
pnj + scale_fill_gradientn(colours = colorRampPalette(brewer.pal(11,"RdYlBu"))(11)) # red to blue  
pnj <- pnj + scale_fill_gradientn(colours = colorRampPalette(rev(brewer.pal(11,"RdYlBu")))(11)) # REV to adjust scale from blue to red 
pnj + aes(y = 300) # make it warming stripes
pnj + aes(x = as.factor(Month), y = 300) # to avoid missing stripes, make month a factor 

# Add labels
pnj + aes(x = as.factor(Month), y = 300) + theme_minimal() + theme(axis.text = element_blank(), axis.title = element_text(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position='right') + labs (x= "Month", y = "Incidents", title = "New Jersey Incidents of Hate Crime", subtitle = "Monthly from 1991 - 2017", caption = "(based on data from the FBI)")

#Make it into a petal circle 
p + coord_polar(theta = "x") 
pnj + coord_polar(theta = "x") 


```

## Monthly ridges
Tried to make this work but the program required the "missing aesthetics: y"

```{r}

p = ggplot(hatedata, aes(x = incident, y = Month, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  labs(title = 'National monthly rates of hate crimes') +
  scale_fill_gradientn(colours = colorRampPalette(rev(brewer.pal(9,"RdYlBu")))(100))
p 

pnj = ggplot(hatedataNJ, aes(x = incident, y = Month, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  labs(title = 'New Jersey monthly rates of hate crimes ') +
  scale_fill_gradientn(colours = colorRampPalette(rev(brewer.pal(9,"RdYlBu")))(100))
pnj
```

